An example cleaning script

### 2025-03-18 code/eda-code/eda.qmd
changes

```{r}
mydata %>% 
  dplyr::filter(food == "Yes") %>%
  count(`Isolation source`) %>%
  top_n(10, n) %>%  # Keep only the top 10 sources
  ggplot(aes(x = reorder(`Isolation source`, -n), y = n)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Top 10 Food 'Yes' Isolation Sources", 
       x = "Isolation Source", 
       y = "Count of Yes") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))
```
#Seeing a few issues with my data cleaning, I will go back to processing file and fix it up

>>>>

```{r}
mydata %>% 
  dplyr::filter(food == "Yes") %>%
  count(`Isolation source`) %>%
  top_n(10, n) %>%  # Keep only the top 10 sources
  ggplot(aes(x = reorder(`Isolation source`, -n), y = n)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Top 10 Food 'Yes' Isolation Sources", 
       x = "Isolation Source", 
       y = "Count of Yes") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))
```
#We see here that the most food isolates for listeria monocytogenes are found in Ready To Eat products, that is a term used by Food Safety and Inspection Service to describe any food products that do not require cooking or sterilization to safely consume.

### 2025-03-18 code/processing-code/processingfile-v1.qmd
changes

```{r}
lmono6 <- lmono5 %>%
  mutate(food = case_when(
    `Isolation type` %in% c("environmental/other", "Not Sliced-Pork", "Sliced-Pork", 
                            "Not Sliced-Turkey", "Sausage Products-Pork", 
                            "Salad/Spread/Pate-Pork", "Salad/Spread/Pate-Chicken", 
                            "Meat/Nonmeat Combination-Beef", "Meat/Nonmeat Combination", 
                            "Sliced-Beef", "Sliced-Turkey", "Not Sliced-Chicken", 
                            "Meat/Nonmeat Combination-Other Poultry", 
                            "Meat/Nonmeat Combination-Combination species", "raw", 
                            "Meat/Nonmeat Combination-Chicken", 
                            "Sausage Products-Unknown species", 
                            "Meat/Nonmeat Combination-Turkey", 
                            "Sausage Products-Beef") & 
    !`Isolation source` %in% c("environmental swab", "environmental swab sponge floor", 
                               "condensate tray", "squeegee", "chain", "wrench", 
                               "underhopper", "environmental swab sponge non food contact surface", 
                               "environmental swab sponge pallet in cooler", 
                               "environmental swab of shelf", 
                               "Environmental Isolate, sponge swab","environmental swab sponge") ~ "Yes",
    TRUE ~ "No"
  ))

>>>>

```{r}
lmono6 <- lmono5 %>%
  mutate(food = case_when(
    `Isolation type` %in% c("environmental/other", "Not Sliced-Pork", "Sliced-Pork", 
                            "Not Sliced-Turkey", "Sausage Products-Pork", 
                            "Salad/Spread/Pate-Pork", "Salad/Spread/Pate-Chicken", 
                            "Meat/Nonmeat Combination-Beef", "Meat/Nonmeat Combination", 
                            "Sliced-Beef", "Sliced-Turkey", "Not Sliced-Chicken", 
                            "Meat/Nonmeat Combination-Other Poultry", 
                            "Meat/Nonmeat Combination-Combination species", "raw", 
                            "Meat/Nonmeat Combination-Chicken", 
                            "Sausage Products-Unknown species", 
                            "Meat/Nonmeat Combination-Turkey", 
                            "Sausage Products-Beef") & 
    !`Isolation source` %in% c("environmental swab", "environmental swab sponge floor", 
                               "condensate tray", "squeegee", "chain", "wrench", 
                               "underhopper", "environmental swab sponge non food contact surface", 
                               "environmental swab sponge pallet in cooler", 
                               "environmental swab of shelf", 
                               "Environmental Isolate, sponge swab","environmental swab sponge","Environmental-factory","Environmental-Wheels on trolley for spinner basket","Environmental-Drain basin between retort 1 and 2","bottom of sliding door in pasta room to blast","bottom of sliding door in pasta room to blast","Environmental-Bottom of sliding door in pasta room to blast","	
environmental swab from drain","environmental", "water sediment","environmental: non-food-contact surface", "soil","manure", "Environmental: food-contact surface","environment","swab") ~ "Yes",
    TRUE ~ "No"
  ))
